% -*-latex-*-
\title{Conflicting Frames: The Dispute Over the Meaning of \emph{Rolezinhos} in Brazilian Media}
\author{Alexandre A. Goncalves}
\department{Department of Comparative Media Studies}
\degree{Master of Science in Comparative Media Studies}
\degreemonth{September}
\degreeyear{2014}
\thesisdate{July 27, 2014}
\copyrightnoticetext{\copyright2014 Alexandre A. Goncalves. All rights reserved.

The author hereby grants to MIT permission to reproduce and to distribute publicly paper and electronic copies of this thesis document in whole or in part in any medium now known or hereafter created.}
\supervisor{Ethan Zuckerman}{Director, Center for Civic Media}
\chairman{T.L. Taylor}{Associate Professor of Comparative Media\\Director of Graduate Studies, CMS}
\maketitle
\cleardoublepage
\begin{abstractpage}
\input{abstract}
\end{abstractpage}
\cleardoublepage
\section*{Acknowledgments}
Once I learned that the English words \enquote{think} and \enquote{thank} share the same origin. It makes sense. To thank is to bring to mind the gifts we received from others. Now, at the end of my stint at MIT, there is a multitude in my mind.

I would like to start by thanking my advisor Ethan for his guidance, confidence, and friendship. I will never be able to repay you for your generous support.

Fox Harrell is certainly one of the brightest and kindest people I have ever met. It is an honor to have been your student and teaching assistant. My chats with Jim Paradis during hikes in New Hampshire represent the most pleasant and restful memories from these two years. Thank you, Jim!

Ed Schiappa has amazed me by his cheerful solicitude for the CMS students. I hope I will follow in your footsteps when I become a teacher. Ed Barrett and Frank Bentley also inspired me when I assisted their class on mobile technology, a model of a challenging and effective course.

Thanks to Sasha Costanza-Chock I was able to connect to the Brazilian community in Boston.

I am also grateful for William Uricchio and Heather Hendershot's detailed comments on my texts---including the tentative thesis proposal and historical chapter that I turned in as assignments for their courses.

My classmates Julie, Edu, Denise, Jason, Ling, Rodrigo, and Erica have been the best fellow travelers one might desire for this kind of journey. I wish to see you succeed in all your dreams.

The 13ers and 15ers 

By naming a few, I express my gratitude to the entire CMS community.

From the Berkman Center, I cannot forget Hal Roberts and David Larochelle, the talented duo behind the Media Cloud code. In the Civic Media side, Rahul Bhargava, Ed Platt, and Erhardt Graeff also lent support and time to help me.

Ben Peters

Jorge Ale
Kevin Majeres, Patricio Fernandez, and all my other friends in Cambridge.

Chelsea and Heather

Leonardo Freitas

Carlos Alberto Di Franco

Alexis and Jude
Civic Media

Brazilian community in Boston.
Thalita Dias

Lorrie
Shannon
Nicole

Caras aqui no Brasil: Ronaldo Lemos, Caff, Renato, Flávio, Dalben

Fernando.
% -*-latex-*-